\documentclass[11pt]{letter}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[margin=1in]{geometry}
\usepackage{microtype}
\usepackage{parskip}

\signature{Egberto F. Selerio Jr.\\
Engineering Research and Development for Technology\\
University of San Carlos, Philippines\\
\texttt{egbertoselerio@gmail.com}}

\address{Egberto F. Selerio Jr.\\
Engineering Research and Development for Technology\\
University of San Carlos, Philippines}

\date{\today}

\begin{document}

\begin{letter}{E. N. Pistikopoulos, PhD FIChemE FAIChE FREng\\
Editor-in-Chief\\
\textit{Computers \& Chemical Engineering}\\
Texas A\&M University\\
College Station, Texas, United States}

\opening{Dear Professor Pistikopoulos,}

Please consider our original research article, ``Model-Agnostic Enforcement of Mass Conservation and Non-Negativity in Statistical Surrogates of Activated Sludge Processes,'' for publication in \textit{Computers \& Chemical Engineering}.

The article addresses a consequential limitation of data-driven process models: accurate predictions are not necessarily physically admissible. We transfer a positivity-preserving conservative projection to steady-state activated-sludge surrogate modeling and evaluate its ability to enforce mass conservation and componentwise non-negativity on complete 20-component effluent predictions. Because the correction is applied after prediction, it provides a common physical safeguard for heterogeneous models without retraining or redesigning them.

The systematic benchmark spans 13 statistical learning methods and examines predictive accuracy, data scarcity, computational cost, and extrapolation beyond the training envelope. Importantly, the analysis distinguishes the guaranteed physical properties of the corrected outputs from the empirical accuracy of the underlying learner. This combination of process modeling, machine learning, and constraint enforcement should be directly relevant to the journal's readership in modeling and simulation, intelligent systems, and process systems engineering, with practical implications for surrogate-assisted screening, optimization, and monitoring of environmental processes.

The manuscript is original, has not been published previously, and is not under consideration elsewhere. All authors have approved its submission and declare no competing interests. Thank you for your consideration.

\closing{Sincerely,}

\end{letter}
\end{document}
