\documentclass[a4paper,10pt]{article}

\usepackage[T1]{fontenc}
\usepackage[margin=0.72in]{geometry}
\usepackage{amsmath}
\usepackage{array}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{microtype}
\usepackage{placeins}
\usepackage{xcolor}
\usepackage{url}

\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}
\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\title{Supplementary Material\\Model-Agnostic Enforcement of Mass Conservation and Non-Negativity in Statistical Surrogates of Activated Sludge Processes}
\author{}
\date{}

\begin{document}

\maketitle

\section{Scope and Reporting Convention}

This supplement reports detailed results from the completed, article-eligible \texttt{article\_final\_v2} run. The benchmark contains 13 statistical surrogates, eleven nested in-distribution (ID) sample totals, five persistent outer folds, and two separately simulated out-of-distribution (OOD) regimes. Raw and projected predictions always refer to the same observations. No OOD response entered fitting, preprocessing, hyperparameter selection, or normalization.

For component $j$, the reported component nRMSE is
\begin{equation}
\mathrm{nRMSE}_j=
\frac{\sqrt{n^{-1}\sum_{i=1}^{n}(\hat c_{ij}-c_{ij})^2}}
{\sigma_j^{(\mathcal T)}},
\end{equation}
where $\sigma_j^{(\mathcal T)}$ is the population SD of that component in the applicable training reference. For a derived composite $q\in\{\mathrm{COD},\mathrm{TN},\mathrm{TP},\mathrm{TSS}\}$, the requested composite-specific counterpart is
\begin{equation}
\mathrm{nRMSE}_q=
\frac{\sqrt{n^{-1}\sum_{i=1}^{n}(\hat y_{iq}-y_{iq})^2}}
{\sigma_q^{(\mathcal T)}}.
\end{equation}
The ID denominator is estimated on the applicable outer-training partition; the OOD denominator is estimated once on all 10,000 ID targets. Composite nRMSE is a secondary, composite-specific normalization and is not the primary nRMSE obtained by averaging squared standardized errors across all 20 effluent components.

ID values are the arithmetic mean $\pm$ sample SD across five outer folds. Each $\Delta$ is calculated within a paired raw/projected fold before those five differences are summarized. OOD values are single descriptive summaries over 600 cases and therefore have no fold SD. No hypothesis tests, confidence intervals, or inferential rank claims are reported.

\section{Benchmark Contract and Dataset Verification}

The 22 inputs were hydraulic retention time, aeration, and 20 influent concentrations. Each surrogate jointly or componentwise predicted the same 20 effluent concentrations in the fixed order $S_O$, $S_F$, $S_A$, $S_{NH4}$, $S_{NO2}$, $S_{NO3}$, $S_{N2}$, $S_{PO4}$, $S_I$, $S_{ALK}$, $X_I$, $X_S$, $X_H$, $X_{PAO}$, $X_{PP}$, $X_{PHA}$, $X_{AOB}$, $X_{NOB}$, $X_{MeP}$, and $X_{MeOH}$. COD, TN, TP, and TSS were calculated after prediction through the same fixed linear composition map for raw and projected outputs.

The surrogate roster comprised XGBoost, LightGBM, CatBoost, AdaBoost, Random Forest, Extra Trees, SVR, $k$-NN, PLS, Multi-task Elastic Net, Multi-task Lasso, MLP, and TabNet. PLS and the two multi-task linear surrogates form the prespecified Interpretable models category because they expose one global coefficient structure; this label does not imply causal interpretation.

\begin{table}[htbp]
\centering
\caption{Mechanistic dataset generation and invariant verification. The invariant residual is $\|Ac_{out}-Ac_{in}\|_2$.}
\label{tab:s_dataset_validation}
\small
\begin{tabular}{lrrrrr}
\toprule
Dataset & Attempted & Retained & Acceptance (\%) & Maximum residual & Minimum component \\
\midrule
In-distribution & 10,000 & 10,000 & 100.0 & $4.90\times10^{-11}$ & $8.75\times10^{-4}$ \\
Mild out-of-distribution & 600 & 600 & 100.0 & $5.67\times10^{-11}$ & $1.62\times10^{-3}$ \\
Severe out-of-distribution & 600 & 600 & 100.0 & $2.91\times10^{-10}$ & $1.01\times10^{-3}$ \\
\bottomrule
\end{tabular}
\end{table}

The ID accuracy export contains $13\times11\times5\times2=1{,}430$ surrogate--size--fold--prediction rows. Component and composite exports contain the corresponding 20-component and four-composite expansions. Tuning completed five 100-trial outer-fold studies and one separate 100-trial full-data study per surrogate, for 7,800 completed trials. The OOD benchmark used one full-data refit per surrogate and retained 600 mechanistic cases per regime.

\section{Detailed In-Distribution Predictive Performance}

Table~\ref{tab:s_id_component_accuracy} gives the same standardized error scale as the primary main-text metric at the individual-component level. The table contains all 260 surrogate--component combinations at $N=10{,}000$ and preserves raw/projected pairing. Projection lowered mean component nRMSE in 188 combinations (72.3\%) and for a majority of components in 12 of 13 surrogates; AdaBoost was the exception at 9 of 20.

\begingroup
\scriptsize
\setlength{\tabcolsep}{2.2pt}
\input{final_results/table_s_id_component_accuracy.tex}
\endgroup

\clearpage
Table~\ref{tab:s_id_composite_accuracy} reports composite-normalized RMSE for COD, TN, TP, and TSS. Projection lowered TN and TP error for every surrogate but increased COD and TSS error for every surrogate. The corresponding physical-unit MSE, RMSE, MAE, bias, maximum absolute error, and $R^2$ remain available in \path{results/article_final_v2/metrics/id_composite_metrics.csv}.

\begingroup
\scriptsize
\setlength{\tabcolsep}{3pt}
\input{final_results/table_s_id_composite_accuracy.tex}
\endgroup

\section{Mass-Conservation, Non-Negativity, and Positivity Backtracking}

All 750,750 raw ID predictions violated the declared mass-conservation tolerance. Raw non-negativity violations varied by surrogate and sample size. After projection, every ID prediction satisfied mass conservation within tolerance and contained no negative concentration; the maximum projected conservation residual was $2.06\times10^{-13}$.

Positivity backtracking was activated for 2,489 predictions (0.332\%). Random Forest, PLS, and Extra Trees accounted for 2,044 activations (82.1\%) and had among the largest standardized projection displacements. Table~\ref{tab:s_backtracking} distinguishes the cumulative count from the activation rate so that the larger number of evaluated rows at high sample totals is not mistaken for a rising backtracking propensity.

\begingroup
\small
\setlength{\tabcolsep}{4pt}
\input{final_results/table_s_backtracking.tex}
\endgroup

\section{Computational Performance}

Fit time includes fold-local preprocessing and surrogate fitting, but excludes Optuna search, artifact serialization, prediction, and projection. Inference timing used a deterministic 512-row batch, five warm-up calls, and 30 measured calls. Values in Table~\ref{tab:s_timing} are five-fold mean $\pm$ sample SD. Search duration is omitted because each fold-specific study was reused at all eleven nested sizes and therefore is not a size-specific training cost.

\begingroup
\small
\setlength{\tabcolsep}{3.5pt}
\input{final_results/table_s_timing.tex}
\endgroup

\section{Detailed Out-of-Distribution Results}

The mild and severe OOD tables use the same 600 cases before and after projection. Component nRMSE uses the fixed full-ID component SD, and composite nRMSE uses the corresponding full-ID composite SD. Projection lowered aggregate 20-component nRMSE for 11 of 13 surrogates under mild extrapolation and 9 of 13 under severe extrapolation. At component level, it lowered 192 of 260 mild comparisons (73.8\%) and 193 of 260 severe comparisons (74.2\%).

\subsection{Mild Extrapolation}

\begingroup
\scriptsize
\setlength{\tabcolsep}{2.2pt}
\input{final_results/table_s_ood_mild_component_accuracy.tex}
\endgroup

\clearpage
\begingroup
\scriptsize
\setlength{\tabcolsep}{3pt}
\input{final_results/table_s_ood_mild_composite_accuracy.tex}
\endgroup

\subsection{Severe Extrapolation}

\begingroup
\scriptsize
\setlength{\tabcolsep}{2.2pt}
\input{final_results/table_s_ood_severe_component_accuracy.tex}
\endgroup

\clearpage
\begingroup
\scriptsize
\setlength{\tabcolsep}{3pt}
\input{final_results/table_s_ood_severe_composite_accuracy.tex}
\endgroup

Every raw OOD prediction set violated mass conservation. Projection eliminated mass-conservation and non-negativity violations across all 15,600 OOD surrogate predictions, with maximum projected conservation residual $1.72\times10^{-13}$. Positivity backtracking was activated 29 times under mild and 39 times under severe extrapolation.

\section{Full-Data Hyperparameters and Scientific Provenance}

Table~\ref{tab:s_full_hyperparameters} records the selected settings used for the final full-ID refits and OOD evaluation. These settings were selected in separate four-fold, 100-trial studies using only ID data. Complete outer-fold selections, trial histories, sampler states, and study databases are retained under \path{results/article_final_v2/tuning/}.

\begingroup
\scriptsize
\setlength{\tabcolsep}{4pt}
\input{final_results/table_s_full_hyperparameters.tex}
\endgroup

\begin{table}[htbp]
\centering
\caption{Scientific-provenance record for the accepted full benchmark.}
\label{tab:s_provenance}
\small
\begin{tabular}{p{4.1cm}p{10.3cm}}
\toprule
Record & Accepted value \\
\midrule
Run identity & \texttt{article\_final\_v2}; full profile; completed 22 July 2026; article eligible \\
Cross-validation & Five persistent outer folds; four inner folds; 100 trials per study; root seed 42 \\
Processor and memory & Intel64 Family 6 Model 181; 12 physical and 14 logical cores; 33,745,088,512 bytes RAM \\
Operating system and Python & Windows 11 (10.0.26200); Python 3.12.13 \\
Core statistical packages & scikit-learn 1.9.0; XGBoost 3.3.0; LightGBM 4.6.0; CatBoost 1.2.10 \\
Neural-network packages & PyTorch 2.13.0; pytorch-tabnet 4.1.0; CPU execution \\
Numerical and data packages & NumPy 2.4.6; SciPy 1.18.0; pandas 3.0.3; PyArrow 25.0.0 \\
Numeric precision & Double precision for mechanistic states, invariant construction, preprocessing, returned predictions, projection, and metrics; single precision inside TabNet \\
Timing protocol & 512 rows; five warm-up calls; 30 measured calls; fold summaries are mean $\pm$ sample SD \\
Completion manifest & SHA-256 \texttt{887798dcb23148806b388add02b9f38e573cff50c1af08383f07fa4d15943b46} \\
\bottomrule
\end{tabular}
\end{table}

\section{Interpretable-Surrogate Coefficient Artifacts and Machine-Readable Results}

The accepted run exports the standardized full-data input--output coefficient matrices for PLS, Multi-task Elastic Net, and Multi-task Lasso. Each artifact contains all 22 named inputs, all 20 component coefficients, the row-wise $L_2$ norm, and an activity indicator at $10^{-12}$. These values describe fitted standardized associations; they are not physical-unit sensitivities, stable feature-selection claims, or causal effects. The files are located under \path{results/article_final_v2/models/interpretability/}.

The numerical source files used for the main text and this supplement are:
\begin{itemize}
\item \path{metrics/id_fold_metrics.csv} and \path{metrics/id_fold_summary.csv} for aggregate ID accuracy;
\item \path{metrics/id_component_metrics.csv} and \path{metrics/id_composite_metrics.csv} for component and composite ID results;
\item \path{metrics/id_physical_fold.csv} and \path{metrics/id_physical_summary.csv} for mass-conservation, non-negativity, displacement, and backtracking diagnostics;
\item \path{metrics/timing_fold_summary.csv} and \path{timing/} for fitting and repeated latency measurements;
\item \path{metrics/ood_metrics.csv}, \path{metrics/ood_component_metrics.csv}, \path{metrics/ood_composite_metrics.csv}, and \path{metrics/ood_physical_summary.csv} for OOD results; and
\item \path{predictions/id/} and \path{predictions/ood/} for paired sample-level raw and projected predictions.
\end{itemize}
All paths above are relative to \path{results/article_final_v2/}. Derived table and figure source CSVs are retained in \path{article/wip/final_results/} and are reproducible with \path{scripts/build_manuscript_revision_assets.py}.

\end{document}
